---
title: "p8105_hw1_xm2356"
author: "Xinyin Miao (xm2356)"
date: "2025-09-12"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(moderndive)
library(tidyverse)
```

# Problem 1

## Dataset description

```{r}
data("early_january_weather")
```

The dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.
The mean temperature is approximately `r round(mean(early_january_weather$temp, na.rm = TRUE), 2)`.
Variables include: `r paste(names(early_january_weather), collapse = ", ")`.


## Scatterplot
```{r}
p_temp_time <- ggplot(early_january_weather,
                      aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.8) +
  labs(title = "Temperature vs Time (colored by humidity)",
       x = "time_hour",
       y = "Temperature",
       color = "humid") +
  theme_minimal()

p_temp_time
```

## Export the plot
```{r}
ggsave("temp_vs_time.png", plot = p_temp_time, width = 7, height = 4)
```


# Problem 2

## Create a data frame
It involves four types of variables and explore means.

```{r}
set.seed(123)
sample_numeric <-  rnorm(10)
sample_logical <-  sample_numeric > 0
sample_character <-  c("apple", "banana", "fish", "dog", "cat", "man", "woman", "book", "store", "test")
sample_factor <- factor(sample(c("low", "medium", "high"), size = 10, replace = TRUE), levels = c("low", "medium", "high"))

library(tibble)
df_types <- tibble(
  number = sample_numeric, 
  is_positive = sample_logical,
  label = sample_character,
  group = sample_factor
)

df_types
```

## Mean of each variable

```{r}
mean(df_types %>% pull(number))
mean(df_types %>% pull(is_positive))
mean(df_types %>% pull(label))  # Error: character vectors have no mean
mean(df_types %>% pull(group)) # Error: Factors can not be averaged directly

```

## Use as.numeric()

```{r echo = TRUE, results = 'hide', warning = FALSE, message=FALSE}
as.numeric(df_types$is_positive)
as.numeric(df_types$label)
as.numeric(df_types$group)
```

## Explanation

"number" is numeric, so the mean works directly.
"is_positive" is logical, TRUE/FALSE are treated as 1/0, so the mean is the proportion of TRUE.
"label" is character, coercion gives NA.
"group" is a factor, as.numeric() returns the internal level codes, not the labels. To get numeric values of the labels, use as.numeric(as.character(group)).










